\normalfont{disCOVIDer19}

### 2020-04-03 Documentation/article_4.tex
update
updated motivation section

 The first one, which happens to be the biggest human pandemic (in terms of spread just as much as deaths \cite{potter2001history} ), is the Spanish Flu: this name is quite misleading since it was not originated in Spain, though the Spanish press was the first one to talk about it. Indeed, at that time (1918-1920) almost every other country put censorships ........ . The origin of this disease is still under research: some scholars focused on the United States \cite{crosby_2003,barry2004site}, on France \cite{shanks2016no} and on Asia \cite{langford2005did}.  The disease was ascribed to the virus A/H1N1, a subtype of influenza A Virus. The spread of the disease was strongly amplified by the fact that many people lived in very bad hygienic conditions. As one can see from Fig.\ref{SpanishFLU} there were basically three waves: the initial spread, the second wave and the third wave. The second wave, that was the most lethal one, coincides with the end of the war and with the coming back of the troops: the close contact inside the trains, and the diffusion of the troops inside their home villages/cities increased abruptly the contagion rate and then the death rate. Furthermore, this effect was amplified by the fact that a more deadly mutation of the virus became widespread \cite{barry2008cross}. As pointed out by different scholars, such effect was enhanced by the fact thst sick people, which infection potential was higher, were transferred by train to the hospitals. Immediately, the virus spread with a higher rate \cite{wever2014death}. As it is possible to note from Fig. \ref{Mortality_spanish_flu} this second wave was largely more deadly for young people with respect the first wave. Finally, looking at plot \ref{SpanishFLU}, a third wave also manifested itself in the beginning of 1919: here, the mortality was higher with respect to the first one, but lower to the second one. In the end, this pandemic infected 500 million people and killed something between 17 and 50 million \cite{spreeuwenberg2018reassessing}. It is worth nothing that, just to compare, the casualties in the Great War were from 8.8 to 10.7 million among soldiers and 11 million among civilian \cite{haythornthwaite1996world}

\begin{figure}[h]
 \centering
 \includegraphics[width=1\linewidth]{Figures/SpanishFLU.png} 
 \caption{Pneumonia and Influenza mortality from 1910-1922 in London: note the three peaks due to the spanish flu. The variation of the diffusion rate R$_{0}$ is reported in the box. Source: \cite{he2011mechanistic} }
 \label{SpanishFLU}
\end{figure}

\begin{figure}[h]
 \centering
 \includegraphics[width=0.6\linewidth]{Figures/Mortality_spanish_flu.png} 
 \caption{Mortality range distribution for the two waves of Spanish flu: note that the second one was more severe for the young people. Source: \cite{taubenberger20061918}}
 \label{Mortality_spanish_flu}
\end{figure}

@@

\begin{figure}[h]
 \centering
 \includegraphics[width=0.8\linewidth]{Figures/Mixing_Flu.jpg} 
 \caption{Wang-ShickRyu diagram for the explanantion of the Spanish, Asian flu and Honk-Kong flu . Source: \cite{RYU2017195}}
 \label{Mixing_Flu}
\end{figure}

\paragraph{Motivation} \label{Motivation}
In order to have a better understanding about the spread and the effects on the population of the COVID19 disease, we propose a ShinyApp, called disCOVIDer19, that gathers and present the data, chiefly coming from Protezione Civile \cite{protezionecivile+git}, of the numbers of new and cumulative infections, hospitalized and intensive care patients, deaths and recoveries in Italy, its regions and provinces; more details will be discussed in sections \ref{Data Origin} and \ref{Panels}. Furthermore we fitted the cumulative cases for Italy and all its regions and provinces with a logistic curve, in order to get a bird-of-eye view on the trend followed as well on the effectiveness of the restrictive measures imposed by the Italian government. The theoretical foundations of this fit ,as well how to manage and use it, will be discussed in the section \ref{Theoretical Background}. Beside the logistic curve, we also considered a statistical approach, largely diffused among economists, that considers the number of cases in each day as a time series: on this basis we were able to make use of a particular tool, which will be briefly introduced in the section \ref{Theoretical Background}, that allowed us to  make a further forecast  about future cases.  This latter becomes useful in case of large deviations from logistic distribution.

>>>>

 The first one, which happens to be the biggest human pandemic (in terms of spread just as much as deaths \cite{potter2001history} ), is the Spanish Flu: this name is quite misleading since it was not originated in Spain, though the Spanish press was the first one to talk about it. Indeed, at that time (1917-1918) almost every other country was involved in the Great War and consequently the censorship was applied to the press . The origin of this disease is still under research: some scholars focused on the United States \cite{crosby_2003,barry2004site}, on France \cite{shanks2016no} and on Asia \cite{langford2005did}.  The disease was ascribed to the virus A/H1N1, a subtype of influenza A Virus. The spread of the disease was strongly amplified by the fact that many people lived in very bad hygienic conditions. As one can see from Fig.\ref{SpanishFLU} there were basically three waves: the initial spread, the second wave and the third wave. The second wave, that was the most lethal one, coincides with the end of the war and with the coming back of the troops: the close contact inside the trains, and the diffusion of the troops inside their home villages/cities increased abruptly the contagion rate and then the death rate. Furthermore, this effect was amplified by the fact that a more deadly mutation of the virus became widespread \cite{barry2008cross}. As pointed out by different scholars, such effect was enhanced by the fact thst sick people, which infection potential was higher, were transferred by train to the hospitals. Immediately, the virus spread with a higher rate \cite{wever2014death}. As it is possible to note from Fig. \ref{Mortality_spanish_flu} this second wave was largely more deadly for young people with respect the first wave. Finally, looking at plot \ref{SpanishFLU}, a third wave also manifested itself in the beginning of 1919: here, the mortality was higher with respect to the first one, but lower to the second one. In the end, this pandemic infected 500 million people and killed something between 17 and 50 million \cite{spreeuwenberg2018reassessing}. It is worth nothing that, just to compare, the casualties in the Great War were from 8.8 to 10.7 million among soldiers and 11 million among civilian \cite{haythornthwaite1996world}

\begin{figure}[h]
 \centering
 \includegraphics[width=1\linewidth]{Figures/SpanishFLU.png} 
 \caption{Pneumonia and Influenza mortality from 1910-1922 in London: note the three peaks due to the spanish flu. The variation of the diffusion rate R$_{0}$ is reported in the box. Image source: Ref. \cite{he2011mechanistic} }
 \label{SpanishFLU}
\end{figure}

\begin{figure}[h]
 \centering
 \includegraphics[width=0.6\linewidth]{Figures/Mortality_spanish_flu.png} 
 \caption{Mortality range distribution for the two waves of Spanish flu: note that the second one was more severe for the young people. Image source: Ref. \cite{taubenberger20061918}}
 \label{Mortality_spanish_flu}
\end{figure}

@@

\begin{figure}[h]
 \centering
 \includegraphics[width=0.8\linewidth]{Figures/Mixing_Flu.jpg} 
 \caption{Wang-ShickRyu diagram for the explanantion of the Spanish, Asian flu and Honk-Kong flu .  Image source: Ref.  \cite{RYU2017195}}
 \label{Mixing_Flu}
\end{figure}

\paragraph{Motivation} \label{Motivation}
In order to have a thorough insight of the spread of the COVID-19 disease and its effects on italian population, we propose a ShinyApp, called disCOVIDer19, that gathers, presents and analyzes the numbers of new and cumulative infections, hospitalized and intensive care patients, deaths and recoveries, over the whole country and over its individual regions and provinces. A detailed guide to the functioning of the App is available in sections \ref{Data Origin} and \ref{Panels}. As regards the analysis panel, we fitted the cumulative cases with a logistic curve, in order to get a bird-of-eye view on the general and specific trends as well as on the effectiveness of the restrictive measures imposed by the Italian government. The theoretical foundations of this fit and how to manage and use it will be discussed in the section \ref{Theoretical Background} and \ref{Data Analysis}. Beside the logistic curve, we also considered a statistical approach, largely diffused among economists, that considers the number of total cases in each day as a time series: on this basis we were able to make use of a particular tool, which will be briefly introduced in the section \ref{Theoretical Background}, that allowed us to make a further forecast about future cases. This latter becomes useful in case of large deviations from logistic distribution.
Finally, we must remark that our approach is informational: we do limit ourselves to list and comment our findings without claiming rather delicate conclusions, expecially about predictions, which many people hinge on nowadays and many lives too, unfortunately.
